---
title: "Final_Project_IST687_Kangsheng_Pu"
author: "Kangsheng Pu"
date: "11/28/2021"
output: pdf_document
---
## Library, Dataset Loading
This session contains all the necessary libraries to complete the projects
```{r}
library(ggplot2)
library(dplyr)
library(kernlab)
library(caret)
library(stringr)
library(rpart)
library(rpart.plot)
df <- read.csv("https://intro-datascience.s3.us-east-2.amazonaws.com/Resort01.csv")
```

## Data cleaning - Variable transformation

```{r}
df_cleaned <- df
df$Meal <- substring(df$Meal, 0 , 2)
# Removing spaces for all string characters for easier reference

# giving numeric values to meal
df_cleaned <- mutate(df , Meal = ifelse(df$Meal == "SC", 0,
                                      ifelse(df$Meal == "BB", 1,
                                      ifelse(df$Meal == "HB", 2,
                                      ifelse(df$Meal == "FB", 3,NA)))));

df_cleaned$stayedDays <- df_cleaned$StaysInWeekendNights + df_cleaned$StaysInWeekNights

# Remove spaces for depositType
df_cleaned$DepositType <- str_replace_all(df_cleaned$DepositType, " ", "")

# Assigning numeric value to deposit type
df_cleaned <- mutate(df_cleaned , DepositType = 
                       ifelse(df_cleaned$DepositType == "NoDeposit", 0,
                       ifelse(df_cleaned$DepositType == "Refundable", 1,
                       ifelse(df_cleaned$DepositType == "NonRefund", 2,
                                      NA))));

# Making new variable checking whether the assigned room is the same as booked room
df_cleaned$AssignedReservedRoom <- ifelse(df_cleaned$ReservedRoomType == df_cleaned$AssignedRoomType, 1, 0)
# 5 new dummy variables for market segment
df_cleaned$cancellationRate <- ifelse(df_cleaned$PreviousBookingsNotCanceled + df_cleaned$PreviousCancellations == 0 , 0.5, df_cleaned$PreviousCancellations / (df_cleaned$PreviousCancellations + df_cleaned$PreviousBookingsNotCanceled))

df_cleaned <- filter(df_cleaned, (df_cleaned$StaysInWeekNights +
                                 df_cleaned$StaysInWeekendNights) >0 &
                                is.na(df_cleaned$Meal)== FALSE &
                                df_cleaned$Adults >0 )
# Factorize the categorical value
df_cleaned$IsCanceled <- as.factor(df_cleaned$IsCanceled)
df_cleaned$Meal <- factor(df_cleaned$Meal, ordered = TRUE, levels = c(0,1,2,3))
df_cleaned$IsRepeatedGuest <- as.factor(df_cleaned$IsRepeatedGuest)
df_cleaned$AssignedReservedRoom <- as.factor(df_cleaned$AssignedReservedRoom)
df_cleaned$DepositType <- factor(df_cleaned$DepositType, ordered = TRUE, levels = c(0,1,2))
df_cleaned$CustomerType <- as.factor(df_cleaned$CustomerType)
df_cleaned$MarketSegment <- as.factor(df_cleaned$MarketSegment)
```

## Data Cleaning - Outliers Removal
```{r}
numericColumns <- c("LeadTime", "StaysInWeekendNights", "StaysInWeekNights",
                    "Adults","PreviousCancellations",
                    "PreviousBookingsNotCanceled", "BookingChanges", 
                    "RequiredCarParkingSpaces"  ,  "TotalOfSpecialRequests" )
attach(df_cleaned)
par(mfrow=c(4,3))
for (eachColumn in numericColumns)
{
  boxplot(df_cleaned[,eachColumn], horizontal = TRUE,xlab = eachColumn)
}
attach(df_cleaned)
par(mfrow=c(4,3))
for (eachColumn in numericColumns)
{
  # Z score outlier removal
  df_cleaned <- filter(df_cleaned , abs((df_cleaned[,eachColumn]- mean(df_cleaned[,eachColumn]))/sd(df_cleaned[,eachColumn]))<3)
  boxplot(df_cleaned[,eachColumn], horizontal = TRUE,xlab = eachColumn)
}

df_cleaned <- filter(df_cleaned, df_cleaned$Children <=2)
```

## Exploratory Data Analysis: Logistic Regression

## Selecting the signficant predictors.
```{r, warning=FALSE}
# Stepwise selection Adding
# Those are the candidates of the possible explanatory variables
candidates <- c("StaysInWeekendNights","StaysInWeekNights" , "Adults",
                "Children", "Babies", "Meal" , "IsRepeatedGuest", "BookingChanges",
                "DepositType", "CustomerType" , "RequiredCarParkingSpaces", 
                "TotalOfSpecialRequests", "AssignedReservedRoom",
                "cancellationRate","PreviousCancellations", 
                "PreviousBookingsNotCanceled", "MarketSegment")
trainList <- createDataPartition(y=df_cleaned$IsCanceled,p=.80,list=FALSE)
trainData <- df_cleaned[trainList,]
testData <- df_cleaned[-trainList,]

glmFormulaFirst <- "IsCanceled ~ "
glmFormulaFirst <- paste0(glmFormulaFirst, paste0(candidates, collapse =  " + "))
glmFormulaFirst <- as.formula(glmFormulaFirst)
logitFirst <- glm(glmFormulaFirst, data = trainData, family = "binomial")
summary(logitFirst)
failed <- c("StaysInWeekendNights", "StaysInWeekNights","Meal", "MarketSegment", "Babies", "RequiredCarParkingSpaces"
            ,"PreviousBookingsNotCanceled", "CustomerType")

for(eachColumn in failed)
{
  candidates <- candidates[!candidates %in% eachColumn]
  glmFailFormula <- "IsCanceled ~ "
  glmFailFormula <- paste0(glmFailFormula, paste0(candidates, collapse =  " + "))
  glmFailFormula <- as.formula(glmFailFormula)
  glmFailTest<- glm(glmFailFormula, data = trainData , family = "binomial")
  # print(summary(glmFailTest))
  # The code above is to help me check each stepwise remove summary so I could check where to stop
  # in console. It's commented.
}
```
## The final logistic model

```{r}
glmFormulaFirst <- "IsCanceled ~ "
glmFormulaFirst <- paste0(glmFormulaFirst, paste0(candidates, collapse =  " + "))
glmFormulaFirst <- as.formula(glmFormulaFirst)
logitFirst <- glm(glmFormulaFirst, data = trainData, family = "binomial")
summary(logitFirst)
```

## Loop through to get average accuracy of the logistic Prediction Model.

```{r, warning=FALSE}
possibilities <- c()
# Generating 30 final accuracy values
for (count in 1:30)
{
  resultAccuracy <- 0
  # THe starting variables are a set of dummy variables that represent Market Segments
  resultColumns <- c("LeadTime")
  for(eachVariable in candidates)
  {
    trainList <- createDataPartition(y=df_cleaned$IsCanceled,p=.80,list=FALSE)
    trainData <- df_cleaned[trainList,]
    testData <- df_cleaned[-trainList,]
    glmFormulaNull <- "IsCanceled ~ LeadTime"
    glmFormulaAlter <- paste(glmFormulaNull, " + ", eachVariable)
    glmFormulaNull <- as.formula(glmFormulaNull)
    glmFormulaAlter <- as.formula(glmFormulaAlter)
    mylogit_null <- glm(glmFormulaNull, data = trainData, family = "binomial")
    mylogit_alternative <- glm(glmFormulaAlter, data = trainData, family = "binomial")
    predLogNull <- predict(mylogit_null, testData, type = "response")
    predLogNull <- ifelse(predLogNull > 0.5, 1, 0) 
    predLogAlter <- predict(mylogit_alternative, testData, type = "response")
    predLogAlter <- ifelse(predLogAlter > 0.5, 1, 0)
    accuracyNull <- sum(diag(table(predLogNull,testData$IsCanceled)))/
      sum(table(predLogNull,testData$IsCanceled))
    accuracyAlter <- sum(diag(table(predLogAlter,testData$IsCanceled)))/
      sum(table(predLogAlter,testData$IsCanceled))
    if(accuracyNull < accuracyAlter)
    {
      resultColumns <- append(resultColumns, eachVariable)
      resultAccuracy <- accuracyAlter
    }
    else{
      resultAccuracy <= accuracyNull
    }
  }
  possibilities <- append(possibilities, resultAccuracy)
}

```



## Goodness of fit Test

```{r}
for (eachVariable in resultColumns)
{
  print(eachVariable)
  print(chisq.test( testData[,eachVariable],testData$IsCanceled))
}
```
## SVM Test

```{r}
SVMcandidates <- c("StaysInWeekendNights","StaysInWeekNights" , "Adults",
                "Children", "Babies", "Meal" , "IsRepeatedGuest", "BookingChanges",
                "DepositType", "CustomerType" , "RequiredCarParkingSpaces", 
                "TotalOfSpecialRequests", "AssignedReservedRoom",
                "cancellationRate","PreviousCancellations", 
                "PreviousBookingsNotCanceled", "MarketSegment")
SVMformula <- "IsCanceled ~ "
SVMformula <- paste(SVMformula, paste(SVMcandidates, collapse = " + "))
SVMformula <- as.formula(SVMformula)
trainList <- createDataPartition(y=df_cleaned$IsCanceled,p=.80,list=FALSE)
trainData <- df_cleaned[trainList,]
testData <- df_cleaned[-trainList,]

trctrl <- trainControl(method = "repeatedcv", number= 5, repeats = 3)

svm.model1 <- train(SVMformula, data = trainData,
                    method = "svmRadial",
                    trControl = trctrl,
                    preProcess= c("center", "scale"),
                    tuneLength = 10)



```

## SVM test results
```{r}
predSVM <- predict(svm.model1,testData)
plot(svm.model1)
svm.model1
```
## Important factors
```{r}
varImp(svm.model1)
```

```{r}
confusionSVM <- confusionMatrix(predSVM, testData$IsCanceled)
confusionSVM
```

## Treebag Model
```{r}
cartTree <- rpart(SVMformula, data = df_cleaned)
treeGraph<-prp(cartTree, faclen = 0, cex = 0.8, extra = 1)
treeGraph
```
## Variables Significance
```{r}
varImp(cartTree)
```

## Treebag predictions and accuracy
```{r}
treeResult <- predict(cartTree, newdata=testData, type = "class")
confusionTree <- confusionMatrix(treeResult, testData$IsCanceled)
confusionTree
```
## Not included in the report: restricting data with customers only in PRT
```{r}
# Stepwise selection Adding
# Those are the candidates of the possible explanatory variables
candidates <- c("StaysInWeekendNights","StaysInWeekNights" , "Adults",
                "Children", "Babies", "Meal" , "IsRepeatedGuest", "BookingChanges",
                "DepositType", "CustomerType" , "RequiredCarParkingSpaces", 
                "TotalOfSpecialRequests", "AssignedReservedRoom",
                "cancellationRate","PreviousCancellations", 
                "PreviousBookingsNotCanceled", "MarketSegment")
trainList <- createDataPartition(y=df_cleaned$IsCanceled,p=.80,list=FALSE)
trainData <- filter(df_cleaned[trainList,], Country == "PRT")
testData <- filter(df_cleaned[-trainList,], Country == "PRT")

glmFormulaFirstPRT <- "IsCanceled ~ "
glmFormulaFirstPRT <- paste0(glmFormulaFirstPRT, paste0(candidates, collapse =  " + "))
glmFormulaFirstPRT <- as.formula(glmFormulaFirstPRT)
logitFirstPRT <- glm(glmFormulaFirstPRT, data = trainData, family = "binomial")
summary(logitFirstPRT)
failed <- c("StaysInWeekendNights", "StaysInWeekNights","Meal", "MarketSegment", "Babies", "RequiredCarParkingSpaces"
            ,"PreviousBookingsNotCanceled", "CustomerType")

for(eachColumn in failed)
{
  candidates <- candidates[!candidates %in% eachColumn]
  glmFailFormula <- "IsCanceled ~ "
  glmFailFormula <- paste0(glmFailFormula, paste0(candidates, collapse =  " + "))
  glmFailFormula <- as.formula(glmFailFormula)
  glmFailTest<- glm(glmFailFormula, data = trainData , family = "binomial")
  # print(summary(glmFailTest))
  # The code above is to help me check each stepwise remove summary so I could check where to stop
  # in console. It's commented.
}
```


